Comment: Reasoning about Interference Between Units

### 2015-08-04 paper/introduction.tex
Peters edits to introduction.tex

\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
hypothesis testing framework to assess counterfactual causal models of
treatment propagation and spillover across social networks.  This research
note improves the statistical infernce presented in \citet{bowers2013sutva}
(henceforth BFP) by substituting a test statistic based on a sum of squared
residuals and incorporating information about the fixed network for the simple
the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
that they used. The point of this short article is to incrementally improve
the use of BFP's ``reasoning about interference'' approach. We do not offer
results about test statistics for multi-parameter causal models on social
networks here, but instead hope that this research note stimulates further,
and deeper, work on test statistics and sharp hypotheses.

@@

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (proportion treated, blocking structure,
etc.), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc.), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc.). In general,
test statistics should be powerful against relevant alternatives (find
canonical cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about large sample performance of certain classes of test statistics
and BFP repeat his general advice: `` Select a test statistic T that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \citet[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produces an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp hypotheses about multiparameter
settings, in practice one may not learn much from such tests --- or, in the
absence of simulation studies assessing the model and test statistic, one may
be misled. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

We assess three versions of the SSR test statistic versus three versions of
the KS test statistic. The first, describe above, we call the SSR+Design test
statistic because it represents information about how treatment is assigned to
the nodes (in $\bz^T \bS$). The second version of the SSR test statistic only
includes network degree and excludes information about the treatment status of
other nodes (only includes $\b1^T \bS$). And the third version includes only
treatment assignment $\bz$. The top row of figure~\ref{fig:twoD} compares the
power of the SSR+Design test statistic (upper left panel) to versions of this
statistic that either only include fixed node degree (SSR+Degree), or no
information about the network at all (SSR). For each test statistic, we tested
the hypothesis $\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation
test (i.e. we sampled 1000 permutations instead of all of them). We executed that
test 10,000 times for each pair of parameters.  The proportion of $p$-values
from that test less than .05 is plotted in Figure~\ref{fig:twoD}: darker
values show fewer rejections, lighter values record more rejections.  All of
these test statistics are valid --- they reject the true null of
$\tau=.5,\beta=2$ no more than 5\% of the time at $\alpha=.05$ --- the plots
are darkest in the area where the two lines showing the true parameters
intersect. All of the plots have some power to reject non-true alternatives
--- as we can see with the large white areas in all of the plots. However,
only when we add information about the number of treated neighbors to the
SSR+Degree statistic, do we see high power against all alternatives in the
plane.

@@

The bottom row of Figure~\ref{fig:twoD} demonstrates the power of the KS test.
The bottom right hand panel shows the test used in BFP. Again all of the tests
are valid in the sense of rejecting the truth no more than 5\% of the time
when $\alpha=.05$ although all of these tests are conservative: the SSR based
tests rejected the truth roughly 4\% of the 10,000 simulations but the KS
tests rejected the truth roughly 2\% of the time.  The KS+Design and KS+Degree
panels show the power of applying the KS test to residuals from linear models
including network degree only (the +Degree version) or degree and also the
number of treated neighbors (the +Design version). That is, whereas the SSR
panels used the sum of squared residuals after accounting for network degree
and/or number of treated neighbors, the KS+Design and KS+Degree panels apply
the KS test to the raw residuals after adjusting for information about the
design and network (or with no adjustment). These panels show (1) that
inclusion of a quantity from the true model (number of treated neighbors) is
not enough to increase power against all alternatives to the level shown by
the SSR+Design test statistic and (2) that the KS tests and the SSR tests have
different patterns of power --- the KS tests might be less powerful in general
(more darker areas on the plots), but our general sense from the analysis
accords with the general sense about the KS test in the scholarly literature
--- that it is relatively low powered.\footnote{For a short description and a
few citations to the well-known low-power of the KS-test, see
\url{https://asaip.psu.edu/Articles/beware-the-kolmogorov-smirnov-test}}

>>>>

@@

Notice that the test statistic choice matters in this process: the engine of
statistical inference involves summarizing information against the
hypothesized claim, yet different ways to summarize information might be more
or less sensitive to substantively meaningful differences. The statistical
power of a simple test of the sharp null hypothesis of no effects will vary as
a function of the design of the study (proportion treated, blocking structure,
etc.), characteristics of the outcome (continuous, binary, skewed, extreme
points, etc.), and the way that a test statistic summarizes the outcome (does
it compare means, standard deviations, medians, qqplots, etc.). In general,
test statistics should be powerful against relevant alternatives (find
canonical cite). \citet[\S 2.4.4]{rosenbaum:2002} provides more specific
advice about large sample performance of certain classes of test statistics
and BFP repeat his general advice: `` Select a test statistic T that will be
small when the treated and control distributions in the adjusted data \ldots
are similar, and large when the distributions diverge.'' \citet[Proposition 4
and 5, \S 2.9]{rosenbaum:2002} presents results proving that test statistics
with this property (``effect increasing'' test statistics), produces an
unbiased test of the hypothesis of no effects or positive effects when the
positive effects involve one parameter.  Yet, when the models are complex and
may involve increasing effects in the direction of one parameter and
non-linear effects in the direction of another parameter, BFP showed that
sometimes a KS-test will have (1) no power to address the model at all such
that all hypothesized parameters would receive the same implausibility
assessment; (2) or might reject all such hypothesized parameters. Thus,
although in theory one may assess sharp hypotheses about multiparameter
settings, in practice one may not learn much from such tests. BFP thus recommended simulation studies of the operating
characteristics of tests as a piece of their workflow --- because  the theory
justifying simple one-dimensional effect increasing test statistics clearly
did not cover multi-parameter situations like those easily arising from social
network experiments.

@@

We assess three versions of the SSR test statistic versus three versions of
the KS test statistic. The first, describe above, we call the SSR+Design test
statistic because it represents information about how treatment is assigned to
the nodes (in $\bz^T \bS$). The second version of the SSR test statistic only
includes network degree and excludes information about the treatment status of
other nodes (only includes $\bOne^T \bS$). And the third version includes only
treatment assignment $\bz$. The top row of figure~\ref{fig:twoD} compares the
power of the SSR+Design test statistic (upper left panel) to versions of this
statistic that either only include fixed node degree (SSR+Degree), or no
information about the network at all (SSR). For each test statistic, we tested
the hypothesis $\tau=\tau_0,\beta=\tau_0$ by using a simulated permutation
test (i.e. we sampled 1000 permutations instead of all of them). We executed that
test 10,000 times for each pair of parameters.  The proportion of $p$-values
from that test less than .05 is plotted in Figure~\ref{fig:twoD}: darker
values show fewer rejections, lighter values record more rejections.  All of
these test statistics are valid --- they reject the true null of
$\tau=.5,\beta=2$ no more than 5\% of the time at $\alpha=.05$ --- the plots
are darkest in the area where the two lines showing the true parameters
intersect. All of the plots have some power to reject non-true alternatives
--- as we can see with the large white areas in all of the plots. However,
only when we add information about the number of treated neighbors to the
SSR+Degree statistic, do we see high power against all alternatives in the
plane.

@@

The bottom row of Figure~\ref{fig:twoD} demonstrates the power of the KS test.
The bottom right hand panel shows the test used in BFP. Again all of the tests
are valid in the sense of rejecting the truth no more than 5\% of the time
when $\alpha=.05$ although all of these tests are conservative: the SSR based
tests rejected the truth roughly 4\% of the 10,000 simulations but the KS
tests rejected the truth roughly 2\% of the time.  The KS+Design and KS+Degree
panels show the power of applying the KS test to residuals from linear models
including network degree only (the +Degree version) or degree and also the
number of treated neighbors (the +Design version). That is, whereas the SSR
panels used the sum of squared residuals after accounting for network degree
and/or number of treated neighbors, the KS+Design and KS+Degree panels apply
the KS test to the raw residuals after adjusting for information about the
design and network (or with no adjustment). These panels show (1) that
inclusion of a quantity from the true model (number of treated neighbors) is
not enough to increase power against all alternatives to the level shown by
the SSR+Design test statistic and (2) that the KS tests and the SSR tests have
different patterns of power --- the KS tests might be less powerful in general
(more darker areas on the plots).
